````markdown
---
title: "Loan Default Exploration Analysis by Fan Fei"
output:
  pdf_document: default
  html_document: default
---

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(dplyr)


```


```{r echo=FALSE, Load_the_Data}
# Load the loan data and the relevant variable definitions

loan <- read.csv('prosperLoanData.csv')
definition <- read.csv('Prosper Loan Data - Variable Definitions - Sheet1.csv')

```

## Dataset Introduction
This loan data set contains 113,937 loans with 81 variables on each loan,
including loan amount, interest rate, current loan status, borrower income, 
borrower employment status, borrower credit history, the latest payment
information, and other relevant factors.

This analysis will explore the relationship between loan default rate and some 
key attributes.

## Univariate Plots & Analysis
Among all the variables below, firstly I will do some preliminary exploration of
the dataset. 

```{r echo=FALSE, Univariate_Plots}

head(loan)
summary(loan)
str(loan)

```

This dataset consists of 81 variables with 113937 observations.

```{r echo=FALSE, Loan_Status}

# Check loan status distribution and default rate
summary(loan$LoanStatus)

LoanStatuspie <- loan %>% 
  group_by(LoanStatus) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(percentage =`n`/sum(`n`)) %>% 
  arrange(desc(LoanStatus))
ggplot(data = LoanStatuspie) +
  geom_bar(aes(x = "", y=percentage, fill = LoanStatus), stat="identity", width = 1) +
  coord_polar("y", start=0) +
  theme_void() +
  geom_text(aes(x = "", y = cumsum(percentage) - percentage/2, label=scales::percent(percentage)), angle = 45, check_overlap = TRUE)

```

Among all the loan status, 'Current' status occpuies almost a half. 'Completed'
status takes the second position, 33.4%. While 4.4% 'Defaulted' and 10.5% 
'Charged off' loan status being categorized into default rate calcuation, in
total of 15% approximately.

```{r echo=FALSE, Categorize_Delinquency}

Delinquentloan <- subset(loan, !is.na(CurrentDelinquencies) & !is.na(DelinquenciesLast7Years))
  Delinquentloan$CurrentDelinquencies[Delinquentloan$CurrentDelinquencies != 0 ] <- " Current Delinquent"
  Delinquentloan$CurrentDelinquencies[Delinquentloan$CurrentDelinquencies == 0 ] <- "Current Non-Delinquent"
  Delinquentloan$DelinquenciesLast7Years[Delinquentloan$DelinquenciesLast7Years != 0 ] <- " Past 7 year Delinquent"
  Delinquentloan$DelinquenciesLast7Years[Delinquentloan$DelinquenciesLast7Years == 0 ] <- " Past 7 year Non-Delinquent"

```

The primary purpose of this analysis is to explore the delinquency related 
factors. For calculation convenience, I firstly categorize the
"CurrentDelinquencies" into "Current Delinquent" and "Current Non-Delinquent"
through non-zero values and zero value. Same as DelinquenciesLast7Years. Later,
most analysis will use Delinquentloan as dataset.

```{r echo=FALSE, Current_Delinquency_Distribution}
    
Delinquenciespie <- Delinquentloan %>% 
  group_by(CurrentDelinquencies) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(percentage =`n`/sum(`n`)) %>% 
  arrange(desc(CurrentDelinquencies))
ggplot(data = Delinquenciespie) +
  geom_bar(aes(x = "", y = percentage, fill = CurrentDelinquencies), stat="identity", width = 1) +
  coord_polar("y", start=0) +
  theme_void() +
  geom_text(aes(x = "", y = cumsum(percentage) - percentage/2, label=scales::percent(percentage)), angle = 45, check_overlap = TRUE)

```

```{r echo=FALSE, DelinquenciesLast7Years_Distribution}

Delinquenciespie7 <- Delinquentloan %>% 
  group_by(DelinquenciesLast7Years) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(percentage =`n`/sum(`n`)) %>% 
  arrange(desc(DelinquenciesLast7Years))
ggplot(data = Delinquenciespie7) +
  geom_bar(aes(x = "", y = percentage, fill = DelinquenciesLast7Years), stat="identity", width = 1) +
  coord_polar("y", start=0) +
  theme_void() +
  geom_text(aes(x = "", y = cumsum(percentage) - percentage/2, label=scales::percent(percentage)), angle = 45, check_overlap = TRUE)

```

Comparing CurrentDelinquencies and Last7YearDelinquencies, nearly 20% loans are 
currently in the delinquent status and there are 32% loans were recorded 
defaulted at least once in the past 7 years.

```{r echo=FALSE, Income_Distribution}

# reorder income range from low to high of StatedMonthlyIncome
loan$IncomeRange = with(loan, reorder(IncomeRange,StatedMonthlyIncome,mean)) 
levels(loan$IncomeRange)

summary(loan$IncomeRange)

Incomepie <- loan %>% 
  group_by(IncomeRange) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(percentage =`n`/sum(`n`)) %>% 
  arrange(desc(IncomeRange))
ggplot(data = Incomepie) +
  geom_bar(aes(x = "", y=percentage, fill = IncomeRange), stat="identity", width = 1) +
  coord_polar("y", start=0) +
  theme_void() +
  geom_text(aes(x = "", y = cumsum(percentage) - percentage/2, label=scales::percent(percentage)), angle = 45, check_overlap = TRUE) + 
  scale_fill_brewer()

```

Income range include both quantitative and descriptive values.Thus,firstly I
reorder the labels based on the mean of a quantitative factor, the Stated 
Monthly Income. Then, people in the "Not Employed" range is the second lowest
but people in the "Not displayed" range may miss values only. 
Based on the pie chart, people with annual income in the range of
25,000 to 75,000 are around 55% of total borrowers. One thing to notice is
that people with higher income (> 75,000) occupies 30% of total borrowers.

```{r echo=FALSE, Monthly_Loan_Payment}

# Check monthly loan payment distribution
ggplot(data = loan, aes(x = MonthlyLoanPayment)) + 
  geom_histogram(binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000))

# Group monthly loan payment into four buckets and check the percentage distribution
loan$MonthlyLoanPayment.bucket <- cut(loan$MonthlyLoanPayment, c(0,250,500,750,1000))

table(loan$MonthlyLoanPayment.bucket, useNA = 'ifany')

Monthlyloanpaymentpie <- loan %>% 
  group_by(MonthlyLoanPayment.bucket) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(percentage =`n`/sum(`n`)) %>% 
  arrange(desc(MonthlyLoanPayment.bucket))
ggplot(data = Monthlyloanpaymentpie) +
  geom_bar(aes(x = "", y=percentage, fill = MonthlyLoanPayment.bucket), stat="identity", width = 1) +
  coord_polar("y", start=0) +
  theme_void() +
  geom_text(aes(x = "", y = cumsum(percentage) - percentage/2, label=scales::percent(percentage)), angle = 45, check_overlap = TRUE) +
  scale_fill_brewer()

```

The distribution of monthly loan payment is shown as right-tailed with over 50%
of loans less than 250. Nearly 90% of loans fall into the range less than 500.


# Bivariate Plots Section

This section will firstly use chi-square test to explore if some quantitative
variables are related to delinquency, using the dataset "DelinquentLoan".
The quantitative factors include: 

BorrowerRate 
EmploymentStatusDuration 
CreditScoreRangeLower 
CurrentCreditLines 
InquiriesLast6Months 
RevolvingCreditBalance 
BankcardUtilization 
DebtToIncomeRatio 
StatedMonthlyIncome 
MonthlyLoanPayment 

Some quantitative factors may have overlaps. For example, CreditScoreRangeLower
and CreditScoreRangeHigher would reflect a similar relationship to Current
Delinquency but with a slightly different x-scale. Thus, only one of these two
would be used to test if such a relationship does exist.

```{r echo=FALSE, Chisquare_Test1}

readborrowerrate <- table(Delinquentloan$BorrowerRate, Delinquentloan$CurrentDelinquencies)
addmargins(round(prop.table(readborrowerrate,1),2),2)
chisq.test(readborrowerrate)

```

```{r echo=FALSE, Chisquare_Test2}

reademployment <- table(Delinquentloan$EmploymentStatusDuration, Delinquentloan$CurrentDelinquencies)
addmargins(round(prop.table(reademployment,1),2),2)
chisq.test(reademployment)

```

```{r echo=FALSE, Chisquare_Test3}

readcreditscore <- table(Delinquentloan$CreditScoreRangeLower, Delinquentloan$CurrentDelinquencies)
addmargins(round(prop.table(readcreditscore,1),2),2)
chisq.test(readcreditscore)

```

```{r echo=FALSE, Chisquare_Test4}

readcreditlines <- table(Delinquentloan$CurrentCreditLines, Delinquentloan$CurrentDelinquencies)
addmargins(round(prop.table(readcreditlines,1),2),2)
chisq.test(readcreditlines)
```

```{r echo=FALSE, Chisquare_Test5}

readinquiries <- table(Delinquentloan$InquiriesLast6Months, Delinquentloan$CurrentDelinquencies)
addmargins(round(prop.table(readinquiries,1),2),2)
chisq.test(readinquiries)

```

```{r echo=FALSE, Chisquare_Test6}

readcreditbalance <- table(Delinquentloan$RevolvingCreditBalance, Delinquentloan$CurrentDelinquencies)
addmargins(round(prop.table(readcreditbalance,1),2),2)
chisq.test(readcreditbalance)

```

```{r echo=FALSE, Chisquare_Test7}

readutilization <- table(Delinquentloan$BankcardUtilization, Delinquentloan$CurrentDelinquencies)
addmargins(round(prop.table(readutilization,1),2),2)
chisq.test(readutilization)

```

```{r echo=FALSE, Chisquare_Test8}

readdebttoincomeratio <- table(Delinquentloan$DebtToIncomeRatio, Delinquentloan$CurrentDelinquencies)
addmargins(round(prop.table(readdebttoincomeratio,1),2),2)
chisq.test(readdebttoincomeratio)

```

```{r echo=FALSE, Chisquare_Test9}

readincome <- table(Delinquentloan$StatedMonthlyIncome, Delinquentloan$CurrentDelinquencies)
addmargins(round(prop.table(readincome,1),2),2)
chisq.test(readincome)

```

```{r echo=FALSE, Chisquare_Test10}

readloanpayment <- table(Delinquentloan$MonthlyLoanPayment, Delinquentloan$CurrentDelinquencies)
addmargins(round(prop.table(readloanpayment,1),2),2)
chisq.test(readloanpayment)

```

Based on all 10 chi-square test results, most factors have p-values < 0.05,
excluding EmploymentStatusDuration and RevolvingCreditBalance. Therefore, the
following plot analysis will focus on the factors with acceptable p-values,
which I can reject the null hypothesis that the factor is independent of the
delinquency.

BorrowerRate - yes plot - not clear
EmploymentStatusDuration - no
CreditScoreRangeLower -yes plot - negative
CurrentCreditLines - yes plot - negative
InquiriesLast6Months - yes plot - positive
RevolvingCreditBalance - no
BankcardUtilization - yes plot - flat within 1.0 and positive after 1.0
DebtToIncomeRatio - yes plot - overall positive
StatedMonthlyIncome - yes plot - overall negative
MonthlyLoanPayment - yes plot - overall negative

```{r echo=FALSE, BorrowerRate_Delinquency}

ggplot(data = loan, aes(x = BorrowerRate, y = CurrentDelinquencies)) +
  geom_line(na.rm = TRUE) +
  scale_x_continuous(limits = c(0.05, 0.30), breaks = seq(0.05, 0.30, 0.05))

Delinquentloan$BorrowerRate.bucket <- cut(Delinquentloan$BorrowerRate, c(0.05,0.1,0.15,0.20,0.25,0.30, 0.35))

table(Delinquentloan$BorrowerRate.bucket, useNA = 'ifany')

BorrowerRateDelinquency <- Delinquentloan %>% 
  group_by(BorrowerRate.bucket,CurrentDelinquencies) %>% 
  count() %>% 
  ungroup() %>% 
  arrange(desc(BorrowerRate.bucket))

ggplot(data = BorrowerRateDelinquency, aes(x = BorrowerRate.bucket, y = n, colour = CurrentDelinquencies)) +       
  geom_line(aes(group = CurrentDelinquencies)) +
  scale_y_continuous(name = "Count", breaks = seq(0, 40000, 5000))

```

```{r echo=FALSE, CreditLines_Delinquency}

ggplot(data = Delinquentloan, aes(x = CurrentCreditLines, fill = CurrentDelinquencies)) +
  geom_histogram(stat = 'count', na.rm = TRUE) +
  scale_x_continuous(limits = c(0, 30), breaks = seq(0, 30, 2)) +
  scale_y_continuous(name = "Count")

```

```{r echo=FALSE, CreditScore_CurrentDelinquency}

ggplot(data = Delinquentloan, aes(x = CreditScoreRangeLower, fill = CurrentDelinquencies)) +
  geom_histogram(stat = 'count', na.rm = TRUE, binwidth = 10) +
  scale_x_continuous(limits = c(500, 850), breaks = seq(500, 850, 50)) +
  scale_y_continuous(name = "Count")

```

```{r echo=FALSE, CreditScore_Last7YearDelinquency}

ggplot(data = Delinquentloan, aes(x = CreditScoreRangeLower, fill = DelinquenciesLast7Years)) +
  geom_histogram(stat = 'count', na.rm = TRUE, binwidth = 10) +
  scale_x_continuous(limits = c(500, 850), breaks = seq(500, 850, 50)) +
  scale_y_continuous(name = "Count")

```

```{r echo=FALSE, Inquiries_CurrentDelinquency}

ggplot(data = Delinquentloan, aes(x = InquiriesLast6Months, fill = CurrentDelinquencies)) +
  geom_histogram(stat = 'count', na.rm = TRUE) +
  scale_x_continuous(limits = c(0,10), breaks = seq(0, 10, 1)) +
  scale_y_continuous(name = "Count", limits = c(0, 30000))

```

```{r echo=FALSE, Inquiries_Last7YearsDelinquency}

ggplot(data = Delinquentloan, aes(x = TotalInquiries, fill = DelinquenciesLast7Years)) +
  geom_histogram(stat = 'count', na.rm = TRUE) +
  scale_x_continuous(limits = c(0,20), breaks = seq(0, 20, 1)) +
  scale_y_continuous(name = "Count")

```

```{r echo=FALSE, BankcardUtilization_Delinquency}

ggplot(data = Delinquentloan, aes(x = BankcardUtilization, colour = CurrentDelinquencies)) +
        geom_line(stat = 'count', na.rm = TRUE) +
        scale_x_continuous(limits = c(0,2), breaks = seq(0, 2, 0.1)) +
        scale_y_continuous(limits = c(0,2000), breaks = seq(0, 1800, 200), name = "Count")

```

```{r echo=FALSE, DebttoIncomeRatio_Delinquency}

ggplot(data = Delinquentloan, aes(x = DebtToIncomeRatio, colour = CurrentDelinquencies)) +
        geom_line(stat = 'count', na.rm = TRUE) +
        scale_x_continuous(limits = c(0,0.9), breaks = seq(0, 0.9, 0.1)) + 
        scale_y_continuous(limits = c(0,3000), name = "Count")

```

```{r echo=FALSE, StatedMonthlyIncome_Delinquency}

IncomeDelinquency <- Delinquentloan %>% 
  group_by(IncomeRange, CurrentDelinquencies) %>% 
  count() %>% 
  ungroup() %>% 
  arrange(desc(IncomeRange))

IncomeRangeOrder = factor(IncomeDelinquency$IncomeRange, levels=c("$0","Not employed","$1-24,999","$25,000-49,999","$50,000-74,999","$75,000-99,999","$100,000+","Not Displayed"))

ggplot(data = IncomeDelinquency, aes(x = IncomeRangeOrder, y = n, colour = CurrentDelinquencies)) +
  geom_line(aes(group = CurrentDelinquencies)) +
  scale_y_continuous(name = "Count", breaks = seq(0, 25000, 2500)) +
  coord_flip()

```

```{r echo=FALSE, MonthlyLoanPayment_Delinquency}

Delinquentloan$MonthlyLoanPayment.bucket <- cut(Delinquentloan$MonthlyLoanPayment, c(0,250,500,750,1000))

table(Delinquentloan$MonthlyLoanPayment.bucket, useNA = 'ifany')

MonthlyloanpaymentDelinquency <- Delinquentloan %>% 
  group_by(MonthlyLoanPayment.bucket,CurrentDelinquencies) %>% 
  count() %>% 
  ungroup() %>% 
  arrange(desc(MonthlyLoanPayment.bucket))

ggplot(data = MonthlyloanpaymentDelinquency, aes(x = MonthlyLoanPayment.bucket, y = n, colour = CurrentDelinquencies)) +       
  geom_line(aes(group = CurrentDelinquencies)) +
  scale_y_continuous(name = "Count", breaks = seq(0, 40000, 5000)) +
  coord_flip()

```

# Bivariate Analysis

Based on all the eight plots above, they somewhat demonstrate relationships with 
current delinquencies/ past 7 years delinquencies times.

First of all, the three discrete variables, credit lines, credit score, and 
inquirie times, I used histogram graph to show their relationship to current
delinquencies times. 
For current credit lines, the more credit lines people have, the less likely 
they would have delinquent loans. With only 1 credit line, the current 
delinquent ratio is nearly 50% of total observations. Bwtween 5-10 credit lines,
the absolute value of observations with delinquent loans approximately maintain 
at the same level while total observations increase. After 10 credit lines, the 
absolute value of observations with delinquent loans decrease while total 
observations decrease also. Overall, the delinquency ratio with credit lines 
variable has a negative relationship.
Credit Score vs. Current Delinquencies and vs. Delinquencies Last 7 Years show
a similar pattern. The higher the credit score is, the lower the delinquency
ratio is. When comparing the two patterns, the 50% ratio line in Credit Score vs
Current Delinquencies is around score at 600. While the 50% ratio line in
Delinquencies Last 7 Years is around score at 650. The delinquency ratio with 
credit score this variable has a negative relationship.
For Inquires vs. Delinquencies, I also employ current and long-term two graphs.
The long-term one (Total Inquiries vs.Delinquencies Last 7 Years) shows a more 
clear relationship.The more inquires people have, they are more likely to have
delinquent loans. The delinquency ratio with total inquires has a positive 
relationship.

Second part is the five continuous variables, borrower rate, bankcard 
utilization, debt-to-income ratio, stated monthly income, and monthly payment.
Since bankcard utilization and debt-to-income ratio fall in a relatively small 
range, around 0 to 2, with 201 and 433 difference values respectively, for these
two I employ a line graph to show the relationship between the utilization ratio
(debt-to-income-ratio) and delinquency ratio. However, both graphs cannot 
demonstrate clear relationships. 
So, looking back to the chi-test table, the bankcard utilization has a flat 
relationship (around 0.2) with delinquency ratio when the utilization is less 
than 1. When bankcard utiliation is more than 1 (meaning people overdraft), the 
delinquency ratio quickly bounce to nearly 0.4 and increase to 1.0 when 
utilization is more than 1.1, which is shown as overlapped lines in the graph.
Also, the debt-to-income ratio vs. delinquency ratio chi-test table tells more
information than the graph. When debt-to-income ratio is less than 1, overall
the delinquency ratio shows a decreasing trend. From 0 to 0.1, the delinquency
ratio is around 0.3 to 0.5. From 0.1 to 0.2, the delinquency ratio is between
0.2 and 0.3. From 0.2 to 0.7, the delinquency ratio is maintained below 0.2.
From 0.7 to 1.0, the ratio increases a little above 0.2. 
Another three variables, borrower rate, monthly income, and loan payment have 
2280, 26870, and 46778 different values, respectively. Thus, I choose to 
transfer them into categorical variables. I use ordered annual income range to 
replace stated monthly income and create borrower rate and loan payment bucket. 
For income range variable, without zero values and not employment observations, 
the overall trend is with income up, the delinquency ratio is decreasing. When 
income range is in between 1 and 24999, the ratio is about 23%. When income 
range is between 25000 and 49999, the ratio is about 27%. In the next income 
ranges, it decreases to about 24%, 22%, and 16.6%, respectively.
For monthly loan payment, the trend is more clear with only four categories. The
more monthly loan payment people need to pay, they would be less likely to have
delinquent loans. With monthly payment bucket under 250, 250 to 500, 500 to 750,
and 750+, the delinquency ratio decreases from about 33.3% to 16% to 12.5% to 0.
For borrower rate, the delinquency ratio is increasing with higher borrower rate
provided. When borrower rate is less than 0.1, the delinquency ratio is lower 
than the overall 20.6% current delinquencies ratio. Then it gradually increases
to 30% when borrower rate goes above 0.3.

# Multivariate Plots Section

```{r echo=FALSE, CreditScore_CreditLines_Delinquencies}

ggplot(aes(x = CreditScoreRangeLower, y = CurrentCreditLines, color = CurrentDelinquencies), data = Delinquentloan) + 
  geom_point(alpha = 1, size = 1, position = 'jitter', na.rm = TRUE) +
  scale_x_continuous(limits = c(500, 850), breaks = seq(500, 850, 50)) +
  scale_y_continuous(limits = c(0,30), breaks = seq(0, 30, 2))

```

```{r echo=FALSE, CreditScore_IncomeRange_Delinquencies}

ggplot(aes(x = CreditScoreRangeLower, y = IncomeRange, color = CurrentDelinquencies), data = Delinquentloan) + 
  geom_jitter(alpha = 1, size = 1, position = 'jitter', na.rm = TRUE) +
  scale_x_continuous(limits = c(500, 850), breaks = seq(500, 850, 50))

```

```{r echo=FALSE, LoanPayment_IncomeRange_Delinquencies}

ggplot(aes(x = MonthlyLoanPayment, y = IncomeRange, color = CurrentDelinquencies), data = Delinquentloan) + 
  geom_jitter(alpha = 1, size = 1, position = 'jitter', na.rm = TRUE) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 250))

```

```{r echo=FALSE, BorrowerRate_DebttoIncomeRatio_Delinquencies}

ggplot(aes(x = BorrowerRate, y = DebtToIncomeRatio, color = CurrentDelinquencies), data = Delinquentloan) + 
  geom_point(alpha = 1, size = 1, position = 'jitter', na.rm = TRUE) +
  scale_y_continuous(limits = c(0,0.9), breaks = seq(0, 0.9, 0.1))

```

```{r echo=FALSE, BankcardUtilization_CreditScore_Delinquencies}

ggplot(aes(x = BankcardUtilization, y = CreditScoreRangeLower, color = CurrentDelinquencies), data = Delinquentloan) + 
  geom_point(alpha = 1, size = 1, position = 'jitter', na.rm = TRUE) +
  scale_x_continuous(limits = c(0, 2), breaks = seq(0, 2, 0.1)) +
  scale_y_continuous(limits = c(500, 850), breaks = seq(500, 850, 20))

```

# Multivariate Analysis

Based on the five graphs above among two variables and the delinquency ratio,
most relationships are aligned with the bivariate analysis. For example, the 
multivariate plot shows Credit Score vs. Credit Lines vs. Delinquencies have
most delinquent loans on the bottom left and least delinquent loans on the upper
right corner, which is consistent with the relationships found in the bivariate
part, the more credit lines and higher credit score people have, they are less
likely to have delinquent loans. ALso, the multivariate plot showing Borrower 
Rate vs. Debt-to-Income Ratio vs. Delinquencies have most delinquent loans on 
the bottom right corner and least delinquent loans on the upper left. It is also
consistent with the observation in the bivariate analysis, the delinquency ratio
has a decreasing trend when debt-to-income increasing. While the Borrower Rate 
bivariate analysis with delinquencies ratio shows a positive relationship. When 
borrower rate is increasing, the delinquent loans ratio is also increasing.
One interesting relationship found in the 2nd and 3rd graph is about the income
range. Though the bivariate plot already demonstrates that when income as "Not
Displayed", the loan default ratio is nearly 50%; the multivariate graph with
monthly loan payment/credit score make it more obvious. The multivariate plot
showing Loan Payment vs. Income Range vs. Delinquencies have most delinquent 
loans concentrated within the "Not Displayed" bracket. 
Another interesting finding is in the 5th graph. There is an area with fewer
observations than other areas. When credit score less than 590 and bank card 
utilization less than 0.5, non-delinquent loans are less than, or at most equal 
to, delinquent loans. Compared to the overall 20% current delinquency ratio,
this area represents a delinquency ratio at least 50%, which also means people
with bank card usage less than 0.5 are less likely to have low credit score,
like < 600. 

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}
    
Delinquenciespie <- Delinquentloan %>% 
  group_by(CurrentDelinquencies) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(percentage =`n`/sum(`n`)) %>% 
  arrange(desc(CurrentDelinquencies))
ggplot(data = Delinquenciespie) +
  geom_bar(aes(x = "", y = percentage, fill = CurrentDelinquencies), stat="identity", width = 1) +
  coord_polar("y", start=0) +
  theme_void() +
  geom_text(aes(x = "", y = cumsum(percentage) - percentage/2, label=scales::percent(percentage)), angle = 45, check_overlap = TRUE) +
  ggtitle("Current Delinquency Ratio")

```

### Description One
This loan dataset contains 113,937 loans with 81 variables on each loan,
including loan amount, interest rate, current loan status, borrower income, 
borrower employment status, borrower credit history, the latest payment
information, and other relevant factors. The most important aspect for loan
analysis is to find out the factors that will contribute to loan delinquencies. 
In the dataset, there are three variables related to loan delinquencies, Current
Delinquencies, Amount Delinquent, and Delinquencies Last 7 Years. All of these 
three variables are numerical data. However, after looking at the dataset and 
categorizing the values of "CurrentDelinquencies" into Current Delinquent (non-
zero values) and Current Non-Delinquent (zero values), the "Current Delinquency
Ratio" pie chart shows only 20% of these loans are non-zero values, 'Current
Delinquent'. Similar result for the Delinquencies Last 7 Years, 32.3%. 
Considering most dependable variables (CurrentDelinquencies / DelinquenciesLast7
years) have zero values, this exploration analysis uses the transformed 
categorical value of "CurrentDelinquencies", rather than the original numerical 
value in the dataset, to explore the relationships between other factors and the
current delinquency ratio. 

### Plot Two
```{r echo=FALSE, Plot_Two_CreditScore}

ggplot(aes(x = CreditScoreRangeLower, y = IncomeRange, color = CurrentDelinquencies), data = Delinquentloan) + 
  geom_jitter(alpha = 1, size = 1, position = 'jitter', na.rm = TRUE) +
  scale_x_continuous(limits = c(500, 850), breaks = seq(500, 850, 50)) + 
  ggtitle("Credit Score vs. Income Range vs. Delinquencies")

```

```{r echo=FALSE, Plot_Two_LoanPayment}

ggplot(aes(x = MonthlyLoanPayment, y = IncomeRange, color = CurrentDelinquencies), data = Delinquentloan) + 
  geom_jitter(alpha = 1, size = 1, position = 'jitter', na.rm = TRUE) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 250)) +
  ggtitle("Loan Payment vs. Income Range vs. Delinquencies")

```

### Description Two
Credit Score and Income Range are two categorical variables with chi-test shown
some relationships with current delinquencies. Combining these two variables
into one jitter point graph, there are some observations found in addition to 
the relationships found in the two bivariate plots. 
Credit Score, individually, has a negative relationship with current delinquency
ratio. Income Range, individually, shows an overall negative trend with current 
delinquency ratio, also.
The combined plot has two interesting blocks. First of all, Credit Score of 600
is a dividing line that roughly seperates left side with more delinquent loans
than non-delinquent and right side with more non-delinquent loans than defaulted
loans. The bivariate histogram between Credit Score and Current Delinquencies
also indicates this 50% dividing line with more orange area than green area on
the left side bars of credit score 600 and more green area than orange area on
the right side bars of credit score 600.
Secondly, the block within "Not Displayed" Income Range and Credit Score less 
than 600 has the most concentrated current delinquencies. Such a concentration
is more obvious in Loan Payment vs. Income Range vs. Delinquencies graph, the
block within "Not Displayed" Income Range and Loan Payment between 0 and 250.
The similarity is the "Not Displayed" Income Range. The bivariate plot
between Income Range and Current Delinquencies has Current Delinquent and 
Current Non-Delinquent crossed at "Not Displayed", meaning a 50% dividing line.
Therefore, credit score of 600 and providable valid income data are two metrics
in deciding the approval of loans.

### Plot Three
```{r echo=FALSE, Plot_Three}

ggplot(aes(x = BankcardUtilization, y = CreditScoreRangeLower, color = CurrentDelinquencies), data = Delinquentloan) + 
  geom_point(alpha = 1, size = 1, position = 'jitter', na.rm = TRUE) +
  scale_x_continuous(limits = c(0, 2), breaks = seq(0, 2, 0.1)) +
  scale_y_continuous(limits = c(500, 850), breaks = seq(500, 850, 20)) +
  ggtitle("Card Utilization vs. Credit Score vs. Delinquencies")

```

### Description Three
In addition to the relationships between these variables and loan delinquency
ratio, the relationships between two independent variables are important, also.
In this plot showing Card Utilization vs. Credit Score vs. Delinquencies, there 
is an area with much fewer observations than other areas, the credit score less 
than 590 and bank card utilization less than 0.5. With more orange jitter points
than green ones, the delinquency ratio in this area is more than 50% dividing 
line. Also, this blank area indicates people are likely to have credit score 
more than 590 with bank card usage less than 0.5. Another dividing line in this 
plot is bankcardutilization of 1.1. Similar as the bivariate plot between 
BankcardUtilization and Current Delinquencies has two lines sharply decreasing
and closely to be overlapped when bankcardutilization more than 1.1; the orange
jitter points are more than green points, especially in the lower credit score 
range, and the total observations are much fewer than the left side of 1.1.

------
  
# Reflection

Overall, this Loan Default Exploration finds out some variables related to the
loan default/delinquencies. The chi-tests, bivariate analysis, and multivariate
plots shows overall relationships as followes:

BorrowerRate - overall positive
CreditScoreRangeLower - negative
CurrentCreditLines - negative
InquiriesLast6Months - positive
BankcardUtilization - flat within 1.0 and positive after 1.0
DebtToIncomeRatio - positive
StatedMonthlyIncome - overall negative
MonthlyLoanPayment - overall negative

However, not all the factors have the same strong relationships with delinquency
ratio. Credit Score, Current Credit Lines, Inquries, Debt-to-Income Ratio have
relatively stronger relationships with delinquency ratio. For borrower rate, 
monthly income, and loan payment, since they are transferred to categorical 
values, we can conclude there are some dividing lines to indicate the overall 
positive/negative trend. BankcardUtilization, however, has most observations
within 1.0, in which has a relatively flat relationship with loan delinquencies.

In addtion to the linear relationship between individual variable and the loan
delinquency ratio, we can also take inter-relationships between variables into
consideration when creating a model. For example, when monthly income not 
provided and credit score less than 600, people are likely to be default. Also, 
people are likely to have credit score more than 590 with bank card usage less 
than 0.5.

Considering most dependable variables have zero values, this analysis uses the 
transformed categorical value of "CurrentDelinquencies", rather than the 
original numerical values in the dataset. So, the analysis cannot indicate how 
many delinquencies a person may have. However, this could be a potential 
exploration on this dataset in the future.
